---
title: "**HW 5 2/29/24**"
author: "Emma Zhou - ez4493 - SDS 315"
output:
  html_document:
  pdf_document:  
    toc: true
    toc_float: yes
---

```{r global_options, echo=FALSE}
knitr::opts_chunk$set(fig.height=4, fig.width=6, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=60))
```

------------------------------------------------------------------------

Link to GitHub: https://github.com/emmazhouu/SDS315_HW5


```{r echo=FALSE, message=FALSE, warning=FALSE}
library(ggplot2)
library(tidyverse)
library(dplyr)
library(mosaic)
library(knitr)
```

## **Q1**
Null hypothesis: Securities trades from Iron Bank are flagged at the same 2.4% baseline rate as that of other traders.

Test statistic: 70 trades were flagged out of 2021

```{r echo=FALSE}
sim_flags <- do(100000) * nflip(2021, 0.024)
ggplot(sim_flags) + geom_histogram(aes(x = nflip), fill = 'darkblue', binwidth = 1)
```

```{r echo=FALSE, results='hide'}
sum(sim_flags >= 70) / 100000
```

The p-value is ~0.002.

Thus, I think the null hypothesis is not plausible because the low p-value implies that the chances of 70 out of 2021 trades or more being flagged by regular chance at the baseline rate of 2.4% is very low. Out of 100,000 simulations of 2021 security trades, only 169 simulations saw 70+ trades flagged, suggesting that the flagging at Iron Bank could be due to something else, not the baseline rate.

## **Q2**
Null hypothesis: Gourmet Bites' rate of health code violations is at the same citywide average of 3%.

Test statistic: 8 out of 50 Gourmet Bites' inspections resulted in health code violations reported
```{r echo=FALSE}
sim_inspect <- do(100000) * nflip(50, 0.03)
ggplot(sim_inspect) + geom_histogram(aes(x = nflip), fill = 'darkblue', binwidth = 1)
```

```{r echo=FALSE, results='hide'}
sum(sim_inspect >= 8) / 100000
```

The p-value is ~0.0001. 

Thus, I think the null hypothesis is not plausible because the very small p-value implies that the chances of 8 out of 50 inspections or more being reported for health code violations at the average rate of 3% due to random issues is very low, even impossible. Out of 100,000 simulations of 50 inspections, only 13 simulations saw 8+ health code violations, suggesting that the higher-than-usual number of health code violations could be due to something else and reflect an actual health code issue, unrelated to the average rate of random issues found typically.

## **Q3**
### **A)**
```{r echo=FALSE, warning=FALSE, results='hide'}
letter_freq <- read.csv('letter_frequencies.csv')
brown_sent <- readLines('brown_sentences.txt')

brown_chi_squared <- numeric(length(brown_sent))

calculate_chi_squared <- function(sentence, freq_table) {
  freq_table$Probability = freq_table$Probability / sum(freq_table$Probability)
  
  # Remove non-letters and convert to uppercase
  clean_sentence = gsub("[^A-Za-z]", "", sentence)
  clean_sentence = toupper(clean_sentence)
  
  # Count the occurrences of each letter in the sentence
  observed_counts = table(factor(strsplit(clean_sentence, "")[[1]], levels = freq_table$Letter))
  
  # Calculate expected counts
  total_letters = sum(observed_counts)
  expected_counts = total_letters * freq_table$Probability
  
  # Chi-squared statistic
  chi_squared_stat = sum((observed_counts - expected_counts)^2 / expected_counts)
  
  return(chi_squared_stat)
}

for (i in seq_along(brown_sent)) {
 brown_chi_squared[i] <- calculate_chi_squared(brown_sent[i], letter_freq)
}

brown_chi_squared <- as_tibble(brown_chi_squared)

ggplot(brown_chi_squared) + geom_histogram(aes(x = value))
```

### **B)**

```{r echo=FALSE, results='hide'}
sentences <- c(
  "She opened the book and started to read the first chapter, eagerly anticipating what might come next.",
  "Despite the heavy rain, they decided to go for a long walk in the park, crossing the main avenue by the fountain in the center.",
  "The museum’s new exhibit features ancient artifacts from various civilizations around the world.",
  "He carefully examined the document, looking for any clues that might help solve the mystery.",
  "The students gathered in the auditorium to listen to the guest speaker’s inspiring lecture.",
  "Feeling vexed after an arduous and zany day at work, she hoped for a peaceful and quiet evening at home, cozying up after a quick dinner with some TV, or maybe a book on her upcoming visit to Auckland.",
  "The chef demonstrated how to prepare a delicious meal using only locally sourced ingredients, focusing mainly on some excellent dinner recipes from Spain.",
  "They watched the sunset from the hilltop, marveling at the beautiful array of colors in the sky.",
  "The committee reviewed the proposal and provided many points of useful feedback to improve the project’s effectiveness.",
  "Despite the challenges faced during the project, the team worked tirelessly to ensure its successful completion, resulting in a product that exceeded everyone’s expectations."
)


# finding chi-squared for test sentences

test_chi_squared <- numeric(10)

for (i in seq_along(sentences)){
test_chi_squared[i] <- calculate_chi_squared(sentences[i], letter_freq)
}

test_chi_squared <- as_tibble(test_chi_squared)

# finding p-values

test_p_values <- numeric(10)

for (i in seq_len(nrow(test_chi_squared))) {
  p_value <- sum(brown_chi_squared >= unlist(test_chi_squared[i,])) / nrow(brown_chi_squared)
  test_p_values[i] <- p_value
}

test_p_values <- round(as_tibble(test_p_values), 3)

test_p_values <- test_p_values %>%
  mutate(sentence_num = c(1:10)) %>%
  select(sentence_num, everything())

test_p_values
```

```{r echo=FALSE}
kable(test_p_values, align = 'c', caption = '<center> **p-values of Test Sentences 1-10** </center>')
```
Sentence 6 was produced by a LLM because its p-value was 0.009, which is noticeably lower than the rest of the p-values. The probability of a reference chi-squared value to be a large or bigger than sentence 6's is 0.009, which is very low. This means the letter frequencies of sentence 6 are abnormal compared to "normal" English sentences from the Brown Corpus.




